### 2024-06-02 2024-5-13/sections/abstract.tex
commit the paper extended by setta2023
update description for ComplexCounting benchmark suite and explanation for experiment results
abstract and intro

Regular expressions (regex for short) and string-length function are widely used in string-manipulating programs. Counting is a frequently used feature in regexes that counts the number of matchings of sub-patterns. 
The state-of-the-art string solvers are incapable of solving string constraints with regex-counting and string-length eﬀiciently, especially when the counting and length bounds are large. 
In this work, we propose an automata-theoretic approach for solving such class of string constraints. 
The main idea is to symbolically model the counting operators by registers in automata instead of unfolding them explicitly, thus alleviating the state explosion problem. Moreover, the string-length function is modeled by a register as well. 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which the off-the-shelf SMT solvers can then solve. 
To improve the performance further, we also propose techniques to reduce the sizes of automata. We implement the algorithms and validate our approach on 48,843 benchmark instances. 
The experimental results show that our approach can solve more instances than the state-of-the-art solvers, at a comparable or faster speed, especially when the counting and length bounds are large.

%\keywords{Rgular Expressions, Counting, Length, Automata, Registers, Linear Integer Arithmetic}
%\end{document}

>>>>

%!TEX root = ../main.tex

Regular expressions (regex for short) and string-length function are widely used in string-manipulating programs. Counting is a frequently used feature in regexes that counts the number of matchings of sub-patterns. 
The state-of-the-art string solvers are incapable of solving string constraints with regex-counting and string-length eﬀiciently, especially when the counting and length bounds are large. 
In this work, we propose an automata-theoretic approach for solving such class of string constraints. 
The main idea is to symbolically model the counting operators by registers in automata instead of unfolding them explicitly, thus alleviating the state explosion problem. Moreover, the string-length function is modeled by a register as well. 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which the off-the-shelf SMT solvers can then solve. 
To improve the performance further, we also propose various optimization techniques.
We implement the algorithms and validate our approach on 49,843 benchmark instances. 
The experimental results show that our approach can solve more instances than the state-of-the-art solvers, at a comparable or faster speed, especially when the counting and length bounds are large or when the counting operators are nested with some other counting operators or complement operators. 
%\denghang{need to be changed before submitting}

### 2024-06-02 2024-5-13/sections/introduction.tex
abstract and intro

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints. 
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers.
  \item We propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
  \item Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. 
 %
  \item Finally, we utilize a collection of benchmark suites comprising 48,843 instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers, especially when the counting and length bounds are large (see Figure~\ref{fig:table_overall_eval} and Table~\ref{tab:large-bound}). For instance, on 1,969 benchmark instances where the counting bounds are greater than or equal to $50$ and the string lengths are required to be beyond $200$, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

@@

\smallskip
\noindent
\emph{Organization.} 
The rest of this paper is structured as follows: Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries. 
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

>>>>

\medskip
\noindent 
\emph{Contribution.} In this work, we focus on the class of string constraints with regular membership and string-length function, where the counting operators may occur (called RECL for brevity). We make the following contributions.
\begin{itemize}
  \item We propose an automata-theoretical approach for solving RECL constraints.
  Our main idea is to represent the counting operators by cost registers in cost-enriched finite automata (CEFA, see Section~\ref{sec:automaton} for the definition), instead of unfolding them explicitly. The string-length function is modeled by cost registers as well. The satisfiability of RECL constraints is reduced to the nonemptiness problem of CEFA w.r.t. a linear integer arithmetic (LIA) formula. According to the results from~\cite{atva2020}, an LIA formula can be computed to represent the potential values of registers in CEFA.
Thus, the nonemptiness of CEFA w.r.t. LIA formulas can be reduced to the satisfiability of LIA formulas, which is then tackled by off-the-shelf SMT solvers. Moreover, we propose techniques to reduce the sizes (i.e. the number of states and transitions) of CEFA, in order to achieve better performance.
%
  \item For complex RECL constraints where the counting operators may be nested with the other counting operators or complement operators, we propose optimization techniques to improve the performance. Furthermore, for the satisfiable RECL constraints where the solutions of the LIA formulas can be obtained, but the models of the original RECL constraint cannot be generated,  we propose optimization techniques to accelerate the model generation process. 
%
%  \item 
%  Combined with the size-reduction techniques mentioned above, the register representation of regex-counting and string-length in CEFA entails an efficient algorithm for solving RECL constraints. 
 %
\item We implement the algorithm on top of OSTRICH, resulting in a string solver called $\ostrichrecl$. Finally, we utilize a collection of benchmark suites comprising 49,843 problem instances in total to evaluate the performance of $\ostrichrecl$. The experiment results show that $\ostrichrecl$ solves the RECL constraints more efficiently than the state-of-the-art string solvers (see Table~\ref{tab:table_overall_eval}-\ref{tab:results_model-gen}). For instance, $\ostrichrecl$ solves 49,638 out of 49,843 instances, while the best state-of-the-art solvers solve only 47,721 instances. We also do experiments to justify some of our technical choices and evaluate the performance of the optimizations. 
%
% 1,000 benchmark instances where the counting and length bounds are large or the counting operators are nested with some other counting operators or complement operators, $\ostrichrecl$ solves at least $278$ more instances than the other solvers, while spending only half or less time per instance on average. 
\end{itemize}

@@

\smallskip
\noindent
\emph{Structure.} 
The rest of this paper is structured as follows: Section~\ref{sec:overview} gives an overview of the approach in this paper. Section~\ref{sec:pre} introduces the preliminaries. 
Section~\ref{sec:recl} presents the syntax and semantics of RECL. 
Section~\ref{sec:automaton} defines CEFA. Section~\ref{sec:algorithm} introduces the algorithm to solve RECL constraints. 
Section~\ref{sec:heuristic} presents some heuristics for improving the performance of solving RECL constraints where the counting operators are nested with some other counting operators or complement operators. Section~\ref{sec:implementation} shows the experiment results. Section~\ref{sec:conclu} concludes this paper.

\smallskip

A preliminary version of this paper has appeared in the proceedings of SETTA 2023~\cite{Denghang2023}. This paper extends the SETTA paper in the following three aspects.
\begin{itemize}
\item To improve the performance of complex RECL constraints where the counting operators are nested with some other counting operators or complement operators, various optimizations are proposed. 
\item To deal with the issue that the satisfiability of LIA formulas can be solved, but models of the original RECL constraints cannot be generated, we propose techniques to optimize the model generation process.  
\item The benchmarks are extended to cove the situations that the counting operators are nested with some other counting operators or complement operators, and the experiments are extended accordingly.  
\end{itemize}
